\documentclass[12pt]{article}

\newcommand{\reporttitle}{A group-theoretical treatment of first-order perturbations of potential boundaries in semiconducting nanocrystals}
\newcommand{\reportauthor}{Michal Horanský}
\newcommand{\reporttype}{CMTH-Vvedensky-3}
\newcommand{\cid}{01881584}


\input{includes}
% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

%\renewcommand{\arraystretch}{1.5}

\begin{document}

%\title{Double Groups and Quantum Dots - Literature Review}%
%\author{Michal Horansky}

%\markboth{M. HORANSKY}%
%{Horansky \MakeLowercase{\textit{et al.}}:}

%\maketitle

\input{titlepage}

\include{sections/1_laypersons_summary}

\section{Abstract}
We investigate the photoluminiscence spectra of InGaAs pyramidal quantum dots with a triangular base under exciton decay. We employ an envelope-function treatment for the exciton wavefunctions, deriving the envelope effective Hamiltonian at the $\Gamma$ point. We review the principles of arguments based on Hamiltonian symmetries in group-theory formalism, specifically the prediction they provide about energy level degeneracies and the selection rules they provide for transitions induced by weak perturbation. We review the claim made by Karlsson \textit{et al.} (2015) that pyramidal quantum dots with triangular base (symmetry group $C_{3v}$) exhibit laser-induced photoluminiscence spectra characteristic of structures with a higher symmetry (symmetry group $D_{3h}$). We propose a theoretical model based on first-order perturbations to symmetry selection rules as a possible mechanism for symmetry elevation, and the condition of localisation within the quantum dot bulk which the model imposes on the wavefunction of an exciton that undergoes symmetry elevation. We claim that exciton complexes with equal charge and equal number of electrons are more likely to undergo symmetry elevation when they are dominated by heavy holes. Because the crystallic lattice in the quantum dot restricts the true Hamiltonian symmetries, we claim that $D_{3h}$ is not a candidate of the elevated symmetry group within our model, and we identify $T_d$ as the only possible candidate symmetry group if no \textit{a priori} symmetry breaking is postulated. We compare the predictions $T_d$ yields to both the data measured by Karlsson \textit{et al.} and a new dataset provided by our colleagues, finding good agreement for a part of the studied sample for several exciton complexes. We review the condition of raising valence band degeneracy at $\Gamma$ which elevating to $T_d$ requires, and which is not treated within our proposed model. We discuss further analysis of elevation to bulk symmetry in nanocrystals, as well as the exciton transformation properties disregarded in our simplistic model.

\pagebreak

\section{Contents}


\renewcommand\contentsname{}

\begingroup
\let\clearpage\relax
\vspace{-1cm} % the removed space. Set as appropriate
\tableofcontents
\endgroup

\pagebreak

\section{Introduction}

Firstly, a historical context of the problem and a brief literature review (namely \cite{karlsson}). Secondly, a brief review of \cite{hexagon} as a \textit{de facto} solution to symmetry elevation, and then motivating the need for a perturbative analysis under strain and imperfect growth, which favour the three pyramid-defined directions out of the six grown. Thirdly, a list of proposed solutions, with the first one, a perturbative model of a low-symmetry Hamiltonian being mathematically explored and quantified.

\pagebreak

\include{sections/5_discussion_of_theory_employed}

\include{sections/6_results}

\include{sections/7_discussion_of_results}

\include{sections/8_conclusions}

\section{Acknowledgements}
The author would like to acknowledge his research partner Zétény Bacsó for his robust, multi-faceted contribution to the project,

their supervisor, Professor Dimitri Vvedensky, for his guidance, direction, and identification of important information sources,

Professor Jing Zhang, for his consultation on the envelope function method and other fundamental properties of quantum dots,

and Ivan Shalashilin, Jacob Edginton, and the author's family for their discussion and support.

\pagebreak

\section{Bibliography}

\begin{thebibliography}{10}

\bibitem{karlsson}
Karlsson, K. F. \textit{et al.} (2015), Spectral signatures of high-symmetry quantum dots and effects of symmetry breaking. \textit{New Journal of Physics}, \textbf{17} 103017

\bibitem{karlsson_2010}
Karlsson, K. F. \textit{et al.} (2010), Fine structure of exciton complexes in high-symmetry quantum dots: Effects of symmetry breaking and symmetry elevation. \textit{Phys. Rev. B}, \textbf{81} 161307(R)

\bibitem{hexagon}
Holsgrove, K. M. \textit{et al} (2022), Towards 3D characterisation of site-controlled InGaAs pyramidal QDs at the nanoscale. \textit{J Mater Sci}, \textbf{57} 16383--16396

\bibitem{bulk_limiting}
Bester, G., Zunger, A. (2005), Cylindrically shaped zinc-blende semiconductor quantum dots do not have cylindrical symmetry: Atomistic symmetry, atomic relaxation, and piezoelectric effects. \textit{Phys. Rev. B}, \textbf{71} 045318

\bibitem{envelope_equation}
Burt, M. G. (1992), The justification for applying the effective-mass approximation to microstructures. \textit{J. Phys.: Condens. Matter}, \textbf{4} 6651. \href{https://doi.org/10.1088/0953-8984/4/32/003}{doi.org/10.1088/0953-8984/4/32/003}

\bibitem{envelope_fundamentals}
Burt, M. G. (1999), Fundamentals of envelope function theory for electronic states and photonic modes in nanostructures. \textit{J. Phys.: Condens. Matter}, \textbf{11} 53

\bibitem{gaas_bonding}
Hsiaw, H. C., Johnson, K. H., Lo, C. F., Adler, D. (1988), Valence and conduction band molecular-orbital topologies and the optical and electrical properties of gallium arsenide and silicon. \textit{J. Non-Cryst. Solids}, \textbf{105} Iss. 1-2 101-106. \href{https://doi.org/10.1016/0022-3093(88)90343-2}{doi.org/10.1016/0022-3093(88)90343-2}

\bibitem{interfaces}
Tomic, S, Vukmirović, N. (2011), Symmetry reduction in multiband Hamiltonians for semiconductor quantum dots: The role of interfaces and higher energy bands. \textit{J. Appl. Phys.}, \textbf{110} Iss. 5 053710. \href{https://doi.org/10.1063/1.3631048}{doi.org/10.1063/1.3631048}

\bibitem{wigner}
Wigner, E. P. (1959), \textit{Group Theory and Its Application to the Quantum Mechanics of Atomic Spectra} (transl. by Griffin, J. J.) (Exp. a. imp. ed.). Academic Press, New York and London. Library of Congress Catalog Card No. 59-10741 (Original work published in 1951)

\bibitem{landau}
Landau, L. D., Lifshitz, E. M. (1977, repr. 1991 w. corrections), \textit{Quantum Mechanics: Non-Relativistic Theory} (transl. by Sykes, J. B., Bell, J. S.) (3rd ed., revised a. enlarged). Pergamon Press, Oxford. ISBN 0-08-020940-8 (Translated from 4th ed. published in 1989)

\bibitem{semiconductor_handbook}
Goldberg, Y. A., Schmidt, N. M. (1996), \textit{Gallium Indium Arsenide ($\text{Ga}_{x}\text{In}_{1-x}\text{As}$)} in \textit{Handbook Series on Semiconductor Parameters: Volume 2: Ternary And Quaternary III-V Compounds} (edited by Levinshtein, M., Rumyantsev, S., Shur, M.). World Scientific, London. ISBN  981-02-1420-0 (Set)

\bibitem{dresselhaus_condensed_matter}
Dresselhaus, M. S., Dresselhaus, G., Jorio, A. (2007), \textit{Group Theory: Application to the Physics of Condensed Matter} (1st ed.). Springer Berlin, Heidelberg. \href{https://doi.org/10.1007/978-3-540-32899-5}{doi.org/10.1007/978-3-540-32899-5}

\bibitem{dresselhaus}
Dresselhaus, M. S. (2002), \textit{Applications of Group Theory to the Physics of Solids}. Massachusetts Institute of Technology.

\bibitem{biedenharn}
Biedenharn, L. C., Louck, J. D. Louck (2009), \textit{Angular Momentum in Quantum Physics: Theory and Application (Encyclopedia of Mathematics and its Applications, Series Number 8)} (illustrated ed.). Addison-Wesley Publishing Company, Advanced Book Program. ISBN 0-201-13507-8

\bibitem{fox}
Fox, M. (2002), \textit{Optical Properties of Solids (Oxford Master Series in Physics)} (1st ed.). Oxford University Press. ISBN 0-19-850613-9

\bibitem{altmann}
Altmann, S. L., Herzig, P. (1994), \textit{Point-Group Theory Tables}. Oxford: Clarendon Press

\end{thebibliography}


\bibliographystyle{unsrt}
% use the following line to create a bibligraphy based on a .bib bibtex file (change to your filename)
%\bibliography{michal_bibliography}





% that's all folks
\end{document}


